\chapter{Introduction}

\section{Methodology}\label{sec:method}
\wannier\ computes maximally-localised Wannier functions (MLWF)
following the method of Marzari and Vanderbilt (MV)~\cite{MV}.
For entangled energy bands, the method of Souza, Marzari and
Vanderbilt (SMV)~\cite{SMV} is used. We introduce briefly the methods
and key definitions here, but full details can be found in the
original papers and in Ref.~\cite{W90}.

First-principles codes typically solve the electronic structure of
periodic materials in terms of Bloch states, $\psi_{n{\bf k}}$. 
These extended states are characterised by a band index $n$ and crystal
momentum ${\bf k}$. An alternative representation can be given in terms
of spatially localised functions known as Wannier functions (WF). The WF 
centred on a lattice site ${\bf R}$, $w_{n{\bf R}}({\bf r})$, 
is written in terms of the set of Bloch states as
\begin{equation}
w_{n{\bf R}}({\bf r})=\frac{V}{(2\pi)^3}\int_{\mathrm{BZ}}
\left[\sum_{m} U^{({\bf k})}_{mn} \psi_{m{\bf k}}({\bf
    r})\right]e^{-\mathrm{i}{\bf k}.{\bf R}} \:\mathrm{d}{\bf k} \ , 
\end{equation}
where $V$ is the unit cell volume, the integral is over the Brillouin
zone (BZ), and $\Uk$ is a unitary matrix that mixes the Bloch 
states at each ${\bf k}$. $\Uk$ is not uniquely defined and different
choices will lead to WF with varying spatial localisations. We define
the spread $\Omega$ of the WF as 
\begin{equation}
\Omega=\sum_n \left[\langle w_{n{\bf 0}}({\bf r})| r^2 | w_{n{\bf
      0}}({\bf r}) \rangle - | \langle w_{n{\bf 0}}({\bf r})| {\bf r}
      | w_{n{\bf 0}}({\bf r}) \rangle |^2 \right].
\end{equation}
The total spread can be decomposed into a gauge invariant term
$\Omega_{\rm I}$ plus a term ${\tilde \Omega}$ that is dependant on the gauge
choice $\Uk$. ${\tilde \Omega}$ can
be further divided into terms diagonal and off-diagonal in the WF basis,
$\Omega_{\rm D}$ and $\Omega_{\rm OD}$,
\begin{equation}
\Omega=\Omega_{\rm I}+{\tilde \Omega}=\Omega_{\rm I}+\Omega_{\rm
  D}+\Omega_{\rm OD} 
\end{equation}
where
\begin{equation}
\Omega_{{\rm I}}=\sum_n \left[\langle w_{n{\bf 0}}({\bf r})| r^2 | w_{n{\bf
      0}}({\bf r}) \rangle - \sum_{{\bf R}m} \left| \langle w_{n{\bf
      R}}({\bf r})| {\bf r} | w_{n{\bf 0}}({\bf r}) \rangle \right| ^2
      \right] 
\end{equation}
\begin{equation}
\Omega_{\rm D}=\sum_n \sum_{{\bf R}\neq{\bf 0}} |\langle w_{n{\bf
    R}}({\bf r})| {\bf r} | w_{n{\bf 0}}({\bf r}) \rangle|^2 
\end{equation}
\begin{equation}
\Omega_{\rm OD}=\sum_{m\neq n} \sum_{{\bf R}} |\langle w_{m{\bf R}}({\bf
  r})| {\bf r} | w_{n{\bf 0}}({\bf r}) \rangle |^2 
\end{equation}
The MV method minimises the gauge dependent spread $\tilde{\Omega}$
with respect the set of $\Uk$ to obtain MLWF.

\wannier\ requires two ingredients from an initial electronic
structure calculation. 
\begin{enumerate}
\item The overlaps between the cell periodic part of the Bloch states
  $|u_{n{\bf k}}\rangle$  
\begin{equation}
M_{mn}^{(\bf{k,b})}=\langle u_{m{\bf k}}|u_{n{\bf k}+{\bf b}}\rangle,
\end{equation}
where the vectors ${\bf b}$, which connect a given k-point with its
neighbours, are determined 
by \wannier\ according to the prescription outlined in Ref.~\cite{MV}.
\item As a starting guess the projection of the Bloch states
  $|\psi_{n\bf{k}}\rangle$ onto trial localised orbitals $|g_{n}\rangle$ 
\begin{equation}
A_{mn}^{(\bf{k})}=\langle \psi_{m{\bf k}}|g_{n}\rangle,
\end{equation}
\end{enumerate}
Note that $\Mkb$, $\Ak$ and $\Uk$ are all small, $N \times
N$ matrices\footnote{Technically, this is true for the case of an
  isolated group of $N$ bands from which we obtain $N$ MLWF. When
  using the disentanglement procedure of Ref.~\cite{SMV}, $\Ak$, for
  example, is a rectangular matrix. See Section~\ref{sec:disentangle}.}
that are independent of the basis set used to obtain the original
Bloch states.

To date, \wannier\ has been used in combination with electronic codes
based on plane-waves and pseudopotentials (norm-conserving and
ultrasoft~\cite{USPP}) as well as mixed basis set techniques such as
FLAPW~\cite{MnO}.

\subsection{Entangled Energy Bands}\label{sec:disentangle}
The above description is sufficient to obtain MLWF for an
isolated set of bands, such as the valence states in an insulator. In
order to obtain MLWF for entangled energy bands we use the
``disentanglement'' procedure introduced in Ref.~\cite{SMV}.

We define an energy window (the ``outer window''). At a given
k-point $\bf{k}$, $N^{({\bf k})}_{{\rm win}}$ states lie within this
energy window. We obtain a set of $N$ Bloch states by
performing a unitary transformation amongst the Bloch states which
fall within the energy window at each k-point: 
 \begin{equation}
| u_{n{\bf k}}^{{\rm opt}}\rangle = \sum_{m\in N^{({\bf k})}_{{\rm win}}}
U^{{\rm dis}({\bf k})}_{mn} | u_{m{\bf k}}\rangle
\end{equation}
where $\bf{U}^{{\rm dis}({\bf k})}$ is a rectangular $N \times N^{({\bf k})}_{{\rm win}}$
 matrix\footnote{As $\bf{U}^{{\rm dis}({\bf k})}$ is a rectangular
 matrix this is a unitary operation in the sense that $(\bf{U}^{{\rm
 dis}({\bf k})})^{\dagger}\bf{U}^{{\rm dis}({\bf k})}=\bf{1}$.}. The
 set of $\bf{U}^{{\rm dis}({\bf k})}$ are obtained by minimising 
 the gauge invariant spread $\Omega_{{\rm I}}$ within the outer energy
 window. The MV procedure can then be used to minimise $\tilde{\Omega}$
 and hence obtain MLWF for this optimal subspace.

It should be noted that the energy bands of this optimal subspace may
not correspond to any of the original energy bands (due to mixing
between states). In order to preserve exactly the properties of a
system in a given energy range (e.g., around the Fermi level) we
introduce a second  energy window. States lying within this inner, or
``frozen'', energy window are included unchanged in the optimal
subspace.

\subsection{Getting Help}
The latest version of \wannier\ and documentation can always
be found at 
{\tt 
\begin{quote}
http://www.wannier.org
\end{quote} }
There is a \wannier\ mailing list for discussing issues in the
development, theory, coding and algorithms pertinent to MLWF.
You can register for this mailing list by following the links at
{\tt
\begin{quote}
http://www.wannier.org/forum.html
\end{quote} }
Finally, many frequently asked questions are answered in
Chapter~\ref{chap:faq}. 

\subsection{Citation}
We ask that you acknowledge the use of \wannier\ in any publications
arising from the use of this code through the following reference
\begin{quote}
[ref] A.~A.~Mostofi, J.~R.~Yates, Y.-S.~Lee, I.~Souza, D.~Vanderbilt
and N.~Marzari,\\
\wannier: A Tool for Obtaining Maximally-Localised Wannier
  Functions,\\ {\it Comput. Phys. Commun.}, submitted (2007); {\tt
  http://arxiv.org/abs/0708.0650}

%%{\tt http://www.wannier.org/}
\end{quote}                                                              

It would also be appropriate to cite the original articles:\\\\
Maximally localized generalized Wannier functions for composite
  energy bands,\\ 
N. Marzari and D. Vanderbilt, {\it Phys. Rev. B} {\bf 56}, 12847 (1997)\\\\
Maximally localized Wannier functions for entangled energy bands,\\
I. Souza, N. Marzari and D. Vanderbilt, {\it Phys. Rev. B} {\bf 65}, 035109 (2001)


\subsection{Credits}
The present release of \wannier\ was written by Arash A. Mostofi
(Imperial College London, UK), Jonathan R. Yates (University of
Cambridge, UK), and Young-Su Lee (KIST, S. Korea). \wannier\ is
based on routines written in 1996-7 for isolated bands by Nicola Marzari
and David Vanderbilt, and for entangled bands by Ivo Souza, Nicola Marzari,
and David Vanderbilt in 2000-1.

Acknowledgements: Stefano de Gironcoli (SISSA, Trieste, Italy) for the
\pwscf\ interface; Michel Posternak (EPFL, Switzerland) for the
original plotting routines. 

\wannier\ \copyright\ 1997-2007 Arash A. Mostofi, Jonathan R. Yates,
Young-Su Lee, Ivo Souza, David Vanderbilt and Nicola Marzari

\subsection{Licence}
All the material in this distribution is free software; you can
redistribute it and/or 
modify it under the terms of the GNU General Public License
as published by the Free Software Foundation; either version 2
of the License, or (at your option) any later version.

This program is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
GNU General Public License for more details.

You should have received a copy of the GNU General Public License
along with this program; if not, write to the Free Software
Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA  02110-1301, USA.


 
