%!TEX root=./user_guide.tex
\chapter{Utilities}\label{ch:utilities}

The \wannier\ code is shipped with a few utility programs that may be
useful in some occasions. In this chapter, we describe their use.

\section{$\tt{kmesh.pl}$}
\label{sec:kmesh}
The \wannier\ code requires the definition of a full Monkhorst--Pack
grid of $k$ points.
In the input file 
the size of this mesh is given by means of the \texttt{mp\_grid} variable. E.g.,
setting
\begin{verbatim}
mp_grid = 4 4 4
\end{verbatim}
tells \wannier\ that we want to use a $4\times 4\times 4$ $k$ grid.

One has then to specify (inside the \texttt{kpoints} block in the 
the \texttt{seedname.win} file) the list of $k$ points of the grid.
Here, the \texttt{kmesh.pl} Perl script becomes 
useful, being able to generate the required list.

The script can be be found in the
\texttt{utility} directory of the \wannier\ distribution.
To use it, simply type:
\begin{verbatim}
./kmesh.pl nx ny nz
\end{verbatim}
where \verb|nx|, \verb|ny| and \verb|nz| define the size of the
Monkhorst--Pack grid that we want to use (for instance, in the above example of the
$4\times 4\times 4$ $k$ grid, 
 \verb|nx|$=$\verb|ny|$=$\verb|nz|$=$4).

This produces on output the list of $k$ points in Quantum
Espresso format, where (apart from a header) the first three columns
of each line are the $k$ coordinates, and the fourth column is the weight
of each $k$ point. This list can be used to create the input file
for the ab-initio \verb|nscf| calculation.

If one wants instead to generate the list of the $k$ coordinates
without the weight (in order to copy and paste the output inside the
\texttt{seedname.win} file), one simply has to provide a fourth
argument on the command line. For instance, for a $4\times 4\times 4$
$k$ grid,
use
\begin{verbatim}
./kmesh.pl 4 4 4 wannier
\end{verbatim}
and then copy the output inside the in the \texttt{kpoints} block in the \texttt{seedname.win} file.

We suggest to always use this utility to generate the $k$ grids. This
allows to
provide the $k$ point coordinates with the accuracy required by
\wannier, and moreover it
makes sure that the $k$ grid used in the ab-initio code and in
\wannier\ are the same.

\section{$\tt{w90chk2chk.x}$\label{sec:w90chk2chk}}
During the calculation of the Wannier functions, \wannier\ produces a
\verb|.chk| file that contains some information to restart the
calculation.

This file is also required by the \postw\ code. 
In particular, the \postw\ code requires at least the \verb|.chk| file, the
\verb|.win| input file, and (almost always) the \verb|.eig|
file. Specific modules may require further files: see the
documentation of each module.

However, the \verb|.chk| file is
written in a machine-dependent format. If one wants to run \wannier\
on a machine, and then continue the calculation with \postw\ on a
different machine (or with \postw\ compiled with a different compiler), the file has to be
converted first in a machine-independent ``formatted'' format on the
first machine, and
then converted back on the second machine.

To this aim, use the \verb|w90chk2chk.x| executable. Note that this
executable is not compiled by default: you can obtain it by executing 
\begin{verbatim}
make w90chk2chk
\end{verbatim}
in the main \wannier\ directory.

A typical use is the following:
\begin{enumerate}
\item Calculate the Wannier functions with \wannier
\item At the end of the calculation you will find a \verb|seedname.chk|
  file. Run (in the folder with this file) the command
\begin{verbatim}
w90chk2chk.x -export seedname
\end{verbatim}
or equivalently
\begin{verbatim}
w90chk2chk.x -u2f seedname
\end{verbatim}
(replacing \verb|seedname| with the seedname of your calculation).

This command reads the \verb|seedname.chk| file and creates a
formatted file  \verb|seedname.chk.fmt| that is safe to be transferred
between different machines.
\item Copy the \verb|seedname.chk.fmt| file (together with the
  \verb|seedname.win| and \verb|seedname.eig| files) on the machine on
  which you want to run \postw.
\item On this second machine (after having compiled
  \verb|w90chk2chk.x|) run
\begin{verbatim}
w90chk2chk.x -import seedname
\end{verbatim}
or equivalently
\begin{verbatim}
w90chk2chk.x -f2u seedname
\end{verbatim}

This command reads the \verb|seedname.chk.fmt| file and creates an
unformatted file  \verb|seedname.chk| ready to be used by \postw.

\item Run the \postw\ code.

\end{enumerate}


\section{$\tt{PL\_assessment}$}
\label{sec:pl_assessment}

The function of this utility is to assess the length of a principal
layer (in the context of a Landauer-Buttiker quantum conductance
calculation) of a periodic system using a calculation on a single unit
cell with a dense k-point mesh.

The utility requires the real-space Hamiltonian in the MLWF basis,
\verb|seedname_hr.dat|. 

The \verb|seedname_hr.dat| file should be copied to a directory containing
executable for the utility. Within that directory, run:

\begin{verbatim}
\$> ./PL_assess.x  nk1 nk2 nk3 num_wann 
\end{verbatim}

where:

 \verb|nk1| is the number of k-points in x-direction
 \verb|nk2| is the number of k-points in y-direction
 \verb|nk3| is the number of k-points in z-direction
 \verb|num_wann| is the number of wannier functions of your system

e.g.,

\begin{verbatim} 
\$> ./PL_assess.x  1 1 20 16
\end{verbatim}

Note that the current implementation only allows for a single k-point
in the direction transverse to the transport direction.

When prompted, enter the seedname.

The programme will return an output file \verb|seedname_pl.dat|, containing four columns
\begin{enumerate}
        \item Unit cell number, $R$
        \item Average 'on-site' matrix element between MLWFs in 
             the home unit cell, and the unit cell $R$ lattice
             vectors away
        \item Standard devaition of the quantity in (2)
        \item Maximum absolute value in (2)
\end{enumerate}

\section{$\tt{w90vdw}$}
\label{sec:w90vdw}

This utility provides an implementation of a method for calculating
van der Waals energies based on the idea of density decomposition via
MLWFs. 

For theoretical details, please see the following publication
and references therein:

Lampros Andrinopoulos, Nicholas D. M. Hine and Arash A. Mostofi, 
``Calculating dispersion interactions using maximally localized
Wannier functions'', \emph{J. Chem. Phys.} \textbf{135}, 154105 (2011).

For further details of this program, please see the documentation
in \verb|utility/w90vdw/doc/| and the related examples in
\verb|utility/w90vdw/examples/|.

\section{$\tt{w90pov}$}
\label{sec:w90pov}

An utility to create Pov files (to render the Wannier functions using
the PovRay utility) is provided inside \verb|utility/w90pov|.

Please refer to the documentation inside \verb|utility/w90pov/doc|
for more information.


\section{$\tt{k\_mapper.py}$}
\label{sec:k_mapper}
The \wannier\ code requires the definition of a full Monkhorst--Pack
grid of $\mathbf{k}$-vectors, which can be obtained by means of the \verb|kmesh.pl| utility.
In order to perform a GW calculation with the Yambo code, you need to perform a nscf calculation on a grid in the irreducible BZ. Moreover, you may need a finer grid to converge the GW calculation than what you need to interpolate the band structure. The \verb|k_mapper.py| tools helps in finding the $\mathbf{k}$-vectors indexes of a full grid needed for interpolation into the reduced grid needed for the GW calculation with Yambo. \newline \newline
Usage:\newline 

\verb|path/k_mapper.py nx ny nz QE_nscf_output|\newline

where \verb|path| is the path of \verb|utility| folder, 
\verb|QE_nscf_output| is the path of the QE nscf output file given to Yambo. 
\section{$\tt{gw2wannier90.py}$}
This utility allows to sort in energy the input data of \verb|wannier90| (e.g. overlap matrices and energy eigenvalues). \verb|gw2wannier90.py| allows to use \verb|wannier90| at the $G_0W_0$ level, where quasi-particle corrections can change the energy ordering of eigenvalues (Some \verb|wannier90| modules require states to be ordered in energy).\newline \newline
Usage:

\verb|path/gw2wannier90.py seedname options|
\newline \newline
where \verb|path| is the path of \verb|utility| folder.\newline\newline

Available options are:
\begin{verbatim}
mmn, amn, spn, unk, uhu, uiu,
spn_formatted, unk_formatted, uhu_formatted, uiu_formatted,
write_formatted
\end{verbatim}

If no options are specified, all the files (\verb|mmn, amn, spn, UNK, uHu, uIu|) are considered.

Binary (unformatted Fortran) files are supported, though not reccommended, 
since they are compiler-dependent.
A safer choice is to use (bigger) formatted files, with options:

\verb|spn_formatted, uiu_formatted, uhu_formatted, unk_formatted|

In default, the output format is the same as the input format.
To generate formatted files with unformatted input, use option:
\verb|write_formatted|
\label{sec:w90aaa}

\section{$\tt{w90spn2spn.x}$\label{sec:w90spn2spn}}
The interface between ab-initio code and \wannier\ 
(e.g. \texttt{pw2wannier90.x}) can produce a
\verb|.spn| file that is used by \postw\ to 
calculate some spin related quantities.

The \verb|.spn| file can be 
written in a machine-dependent or a machine-independent format 
depending on the input parameter \texttt{spn\_formatted} 
(the default is \texttt{false} which means the \verb|.spn| file 
is machine-dependent) of the 
\texttt{pw2wannier90.x}. If a \verb|.spn| file has been generated 
on a machine with machine-dependent format, 
and then one wants to continue the calculation 
with \postw\ on a different machine 
(or with \postw\ compiled with a different compiler), 
the file has to be converted first in a machine-independent 
``formatted'' format on the first machine. 

To this aim, use the \verb|w90spn2spn.x| executable. Note that this
executable is not compiled by default: you can obtain it by executing 
\begin{verbatim}
make w90spn2spn
\end{verbatim}
in the main \wannier\ directory.

A typical use is the following:
\begin{enumerate}
	\item Calculate the \verb|.spn| file, e.g. by \texttt{pw2wannier90.x}
	\item At the end of the calculation you will find a \verb|seedname.spn|
	file. If the file is ``unformatted'', run (in the folder with this file) the command
	\begin{verbatim}
	w90spn2spn.x -export seedname
	\end{verbatim}
	or equivalently
	\begin{verbatim}
	w90spn2spn.x -u2f seedname
	\end{verbatim}
	(replacing \verb|seedname| with the seedname of your calculation).
	
	This command reads the \verb|seedname.spn| file and creates a
	formatted file  \verb|seedname.spn.fmt| that is safe to be transferred
	between different machines.
	\item Copy the \verb|seedname.spn.fmt| file on the machine on
	which you want to run \postw.
	\item On this second machine (after having compiled
	\verb|w90spn2spn.x|) run
	\begin{verbatim}
	w90spn2spn.x -import seedname
	\end{verbatim}
	or equivalently
	\begin{verbatim}
	w90spn2spn.x -f2u seedname
	\end{verbatim}
	
	This command reads the \verb|seedname.spn.fmt| file and creates an
	unformatted file \verb|seedname.spn| ready to be used by \postw.
	
	\item Run the \postw\ code.
	
\end{enumerate}

Note if \texttt{spn\_formatted} is set to \texttt{true} in both 
\texttt{pw2wannier90.x} and \postw\ input files, then the \verb|.spn| file will 
be written and read as ``formatted'', so \texttt{w90spn2spn.x} is not needed. 
However, if an ``unformatted'' \verb|seedname.spn| has been created and 
you do not want to rerun \texttt{pw2wannier90.x}, then 
\texttt{w90spn2spn.x} can be useful. 
Also, once a ``formatted'' \verb|seedname.spn| has been generated, the step 4 can be skipped if 
\texttt{spn\_formatted} is set to \texttt{true} in \postw\ input file \verb|seedname.win|.

\section{$\tt{mb.py}$\label{sec:mb}}
This utility allows to incorporate many-body excitonic effects into the independent-particle linear and quadratic optical responses by means of time-dependent current-density functional theory including the electron-hole attraction \textit{via} a tensorial long-range exchange-correlation kernel, as explained in Ref.~\cite{garcia-goiricelaya-prb23}.

At first order, the utility requires the linear optical response file, {\tt seedname-eps\_ij.dat}, where the $ij$ components of the complex linear optical susceptibility and conductivity and those of the complex optical dielectric tensor within the independent-particle approximation are written as a function of energy. At the microscopic (m) level, the optical dielectric tensor including long-range excitonic effects reads as
\begin{equation}
 \varepsilon^{ab}_{\mathrm{m}}(\omega)=\delta_{ab}+\sum_{c}\left(\delta_{ac}-\frac{\alpha^{ac}_{\mathrm{LRC}}}{4\pi}\right)\chi^{cb}(\omega),
\end{equation}
with the coefficients $\alpha^{ab}_{\mathrm{LRC}}$ governing the screening of the attractive Coulomb-like potential acting as the long-range contribution (LRC) of the tensorial exchange-correlation kernel. At the macroscopic (M) level, the inverse of the optical dielectric tensor including long-range excitonic effects reads as
\begin{equation}
 \varepsilon^{-1,ab}_{\mathrm{M}}(\omega)=\delta_{ab}-\chi_{\mathrm{m}}^{ab}(\omega),
\end{equation}
where the linear optical susceptibility including long-range excitonic effects at the microscopic and macroscopic levels are given, respectively, by
\begin{subequations}
 \begin{equation}
  \chi_{\mathrm{m}}^{ab}(\omega)=\sum_{c}\chi^{ac}(\omega)\varepsilon^{cb}_{\mathrm{m}}(\omega),
 \end{equation}
 \begin{equation}
  \chi_{\mathrm{M}}^{ab}(\omega)=\sum_{c}\chi_{\mathrm{m}}^{ac}(\omega)\varepsilon^{cb}_{\mathrm{M}}(\omega).
 \end{equation}
\end{subequations}
In the first two equations, all magnitudes are evaluated in atomic units in order to ensure optical dielectric tensors are dimensionless. In the last two equations, long-range excitonic effects are included into the optical susceptibility, as it can be also done for the optical conductivity.

As output, the utility writes the $ij$ components of the complex linear optical susceptibility and conductivity and those of the complex optical dielectric tensor including long-range excitonic effects in the files {\tt seedname-eps-mb\_ij.dat}.

At second order, the utility requires the linear optical response file, {\tt seedname-eps\_ij.dat}, and the quadratic optical response file: in the case of shift-current, the file {\tt seedname-sc\_ijk.dat}, and in the case of second-harmonic generation, the file {\tt seedname-shg\_ijk.dat}, where the $ijk$ components of the complex quadratic optical susceptibility and conductivity within the independent-particle approximation are written as a function of energy. The quadratic optical susceptibility tensor including only linear long-range excitonic effects at the microscopic and macroscopic level, respectively, read as
\begin{subequations}
 \begin{equation}
  \chi^{abc}_{\mathrm{m}}(\omega_{12};\omega_{1},\omega_{2})=\sum_{def}\varepsilon^{-1,ad}_{\mathrm{m}}(\omega_{12})\chi^{def}(\omega_{12};\omega_{1},\omega_{2})\varepsilon^{-1,eb}_{\mathrm{m}}(\omega_{1})\varepsilon^{-1,fc}_{\mathrm{m}}(\omega_{2}),
 \end{equation}
 \begin{equation}
  \chi^{abc}_{\mathrm{M}}(\omega_{12};\omega_{1},\omega_{2})=\sum_{def}\varepsilon^{ad}_{\mathrm{M}}(\omega_{12})\chi^{def}_{\mathrm{m}}(\omega_{12};\omega_{1},\omega_{2})\varepsilon^{eb}_{\mathrm{M}}(\omega_{1})\varepsilon^{fc}_{\mathrm{M}}(\omega_{2}),
 \end{equation}
\end{subequations}
where quadratic long-range excitonic effects (\textit{i.e.}, exchange-correlation kernels at second order) are neglected because even their expression is unknown.

As output, the utility writes the $ijk$ components of the quadratic optical susceptibility and conductivity including long-range excitonic effects in the files {\tt seedname-sc-mb\_ijk.dat} and {\tt seedname-shg-mb\_ijk.dat} for the shift current and second-harmonic generation contributions, respectively.

A typical use is the following:

\begin{verbatim}
\$> path/mb.py
\end{verbatim}

where \verb|path| is the utility folder. Once executed:

\begin{enumerate}
 \item The user is asked to write the \verb|seedname| of the calculation
 \item The user is asked to choose an optical process. The valid options are:
 \begin{enumerate}
  \item \verb|eps|: linear optical response
  \item \verb|sc|: shift current contribution to the quadratic optical response
  \item \verb|shg|: second-harmonic-generation contribution to the quadratic optical response
 \end{enumerate}
 \item Depending on the chosen optical process, it is checked if all the necessary files are present and if they have the same energy values in the same energy range. Then, the user is asked to choose the procedure for estimating the long-range excitonic screening coefficients $\alpha^{ab}_{\mathrm{LRC}}$:
 \begin{enumerate}
  \item A non-selfconsistent bootstrap calculation, also known as RPA-bootstrap (RBO)~\cite{PhysRevLett.114.146402,PhysRevLett.115.137402}
  \item A self-consistent bootstrap calculation (BO)~\cite{PhysRevLett.107.186401}
  \item Empirical values
 \end{enumerate}
 \item If the non-selfconsistent or the self-consistent bootstrap procedure is chosen, the user is asked wether the material is metallic. If yes, the user is informed that bootstrap approximations can not be used for metals, and the user must start again using empirical values. On the contrary, if the material is not metallic, bootstrap approximations can be used. Finally, the coefficients $\alpha^{ab}_{\mathrm{LRC}}$ of the used tensorial LRC exchange-correlation kernel are prompted, and the chosen optical response including long-range excitonic effects is written as output.
\end{enumerate}
