\chapter{Files}


\section{seedname.win}
INPUT. The master input file; contains the specification of the system
and any parameters for the run. 

\subsection{Units}

The following are the dimensional quantities that are
specified in the master input file:

\begin{itemize}
\item Direct lattice vectors
\item Positions (of atomic or projection) centres in real space
\item Energy windows
\item Positions of $\mathbf{k}$-points in reciprocal space
\item \verb#zona# and \verb#box-size# (see Section~\ref{sec:proj})
\end{itemize}

Notes:

\begin{itemize}
\item The units (either \verb#ang#
  (default) or \verb#bohr#) can be set in the first line of the blocks \verb#unit_cell_cart# and \verb#atoms_cart#
\item Energy is always in eV.
\item Positions of $\mathbf{k}$-points are always in crystallographic
  coordinates
relative to the reciprocal lattice vectors.
\item \verb#box-size# and \verb#zona# always in Angstrom and
  reciprocal Angstrom, respectively
\item The keyword \verb#length_unit# may be set to \verb#ang#
  (default) or \verb#bohr#, in order to set the units in which the
  quantities in the output file are written.
\end{itemize}

The reciprocal lattice vectors
$\{\mathbf{B}_{1},\mathbf{B}_{2},\mathbf{B}_{3}\}$ are defined in
terms
of the direct lattice vectors
$\{\mathbf{A}_{1},\mathbf{A}_{2},\mathbf{A}_{3}\}$ by the equation

\begin{equation}
\mathbf{B}_{1} = \frac{2\pi}{\Omega}\mathbf{A}_{2}\times\mathbf{A}_{3}
\ \ \ \mathrm{etc.},
\end{equation}

where the cell volume is
$\Omega=\mathbf{A}_{1}\cdot(\mathbf{A}_{2}\times\mathbf{A}_{3})$.

\section{seedname.mmn}
INPUT. See Chapter~\ref{ch:wann-pp}.

\section{seedname.amn}
INPUT. See Chapter~\ref{ch:wann-pp}.

\section{seedname.eig}
INPUT. See Chapter~\ref{ch:wann-pp}.

\section{seedname.nnkp} \label{sec:old-nnkp}
OUTPUT. See Chapter~\ref{ch:wann-pp}.

\section{seedname.wout}
OUTPUT. The master output file.


\section{seedname.chk}
INPUT/OUTPUT. Sufficient information to restart the calculation or enter the
plotting phase.

\section{UKNn.s}
INPUT. Read if \verb#wannier_plot#=\verb#TRUE# and used to plot the Wannier functions. 

The periodic part of the bloch states represented on a regular real space grid,
 indexed by kpoint n and spin s. 
The first line of the file should contain 5 integers; the number of grid points in each
direction (ngx, ngy and ngz), the kpoint number and the total number of bands in
the file. The full file will be read by wannier90 as:

\begin{verbatim}
read(file_unit) ngx,ngy,ngz,ik,nbnd
do loop_b=1,num_bands
  read(file_unit) (r_wvfn(nx,loop_b),nx=1,ngx*ngy*ngz)
end do
\end{verbatim}

The file can be in formatted or unformatted style, this is controlled by the logical keyword \verb#wvfn_formatted#.



