\chapter{Overview}

\section{Methodology}
Wannier90 computes Maximally Localised
Wannier Functions following the method of Marzari and Vanderbilt (MV).\footnote{
{\it Maximally localized generalized Wannier functions for composite energy bands}
N. Marzari and D. Vanderbilt, Phys. Rev. B 56, 12847 (1997)}
For entangled energy bands  the method of
Souza, Marzari and Vanderbilt (SMV)\footnote{{\it Maximally localized
    Wannier functions for entangled energy bands} 
I. Souza, N. Marzari and D. Vanderbilt, Phys. Rev. B 65, 035109 (2002)}
is used.
We briefly introduce the methods and key definitions, but full details
can be found in the original papers. 

First principles codes typically solve the electronic structure of
periodic materials in terms of Bloch states, $\psi_{n{\bf k}}$. 
These extended states are characterised by a band index, n and crystal
momentum, ${\bf k}$. An alternative representation can be given in terms
of spatially localised functions known as Wannier functions (WF). The Wannier function 
centred on a lattice site ${\bf R}$,  $w_{n{\bf R}}({\bf r})$, 
is written in terms of the set of Bloch states as
\begin{equation}
w_{n{\bf R}}({\bf r})=\frac{V}{(2\pi)^3}\int_{BZ} \left[\sum_{m} U^{({\bf
k})}_{mn} \psi_{m{\bf k}}({\bf r})\right]e^{-{\bf k}.{\bf R}} d{\bf k},
\end{equation}
where $U^{({\bf k})}$ is a unitary matrix which mixes the Bloch states at each
${\bf k}$. $U^{({\bf k})}$ is not uniquely defined and different choices
will lead to WF with varying spatial localisations. We define the spread
of the Wannier functions, $\Omega$ 
\begin{equation}
\Omega=\sum_n \left[\langle w_{n{\bf 0}}({\bf r})| r^2 | w_{n{\bf
      0}}({\bf r}) \rangle - | \langle w_{n{\bf 0}}({\bf r})| {\bf r} | w_{n{\bf
      0}}({\bf r}) \rangle |^2 \right].
\end{equation}
The total spread can be decomposed into a gauge invariant term,
$\Omega_{\rm I}$ and a term, ${\tilde \Omega}$ which is dependant on the gauge
choice $U^{({\bf k})}$. ${\tilde \Omega}$ can
be further divided into terms diagonal and off-diagonal in the WF basis,
$\Omega_{\rm D}$ and $\Omega_{\rm OD}$.
\begin{equation}
\Omega=\Omega_{\rm I}+{\tilde \Omega}=\Omega_{\rm I}+\Omega_{\rm D}+\Omega_{\rm OD}
\end{equation}
where
\begin{equation}
\Omega_{{\rm I}}=\sum_n \left[\langle w_{n{\bf 0}}({\bf r})| r^2 | w_{n{\bf
      0}}({\bf r}) \rangle - \sum_{{\bf R}m} \left| \langle w_{n{\bf R}}({\bf r})| {\bf r} | w_{n{\bf
      0}}({\bf r}) \rangle \right| ^2 \right]
\end{equation}
\begin{equation}
\Omega_{\rm D}=\sum_n \sum_{{\bf R}\neq{\bf 0}} |\langle w_{n{\bf R}}({\bf r})| {\bf r} |
w_{n{\bf 0}}({\bf r}) \rangle|^2 
\end{equation}
\begin{equation}
\Omega_{\rm D}=\sum_{m\neq n} \sum_{{\bf R}} |\langle w_{n{\bf R}}({\bf
  r})| {\bf r} |
w_{n{\bf 0}}({\bf r}) \rangle |^2 
\end{equation}
The MV method minimises the gauge dependent spread $\tilde{\Omega}$, with respect the set
of $U^{({\bf k})}$ to obtain Maximally Localised Wannier Functions (MLWF).

The Wannier90 code requires two ingredients from an initial electronic structure calculation.
\begin{enumerate}
\item The overlaps between the cell periodic part of the bloch states $|u_{n{\bf k}}\rangle$ 
\begin{equation}
M_{mn{\bf k}}=\langle u_{m{\bf k}}|u_{n{\bf k}+{\bf b}}\rangle,
\end{equation}
where the vectors ${\bf b}$, which connect a given kpoint with its neighbours, are determined
by the Wannier90 code according to prescription outlined in MV.
\item As a starting guess the projection of the bloch states onto trial localised orbitals 
\begin{equation}
M_{mn{\bf k}}=\langle \psi_{m{\bf k}}|\psi_{n{\bf k}+{\bf b}}\rangle,
\end{equation}
\end{enumerate}
Note that $M_{{\bf k}}$, $A_{{\bf k}}$ and $U^{({\bf k})}$ are all
small, $N_{{\rm wann}}^2$ matrices, independent of the basis set used to
obtain the original Bloch states. To date the Wannier90 code has been
used in combination with electronic codes based on planewaves and
pseudopotentials (norm-conserving and ``ultrasoft'') as well as mixed
basis set techniques such as FLAPW. 

\subsection{Entangled Energy Bands}
The above description is sufficient to obtain Wannier functions for
isolated set of bands, such as the valence states in an insulator. In
order to obtain MLWF for entangled energy bands we use the
``disentanglement'' procedure introduced in SMV.

We define an energy window (the ``outer window''). At a given kpoint 
$N^{{\bf k}}_{{\rm win}}$ states lie within this energy window. We obtain a set of
$N_{{\rm wann}}$ Bloch states by performing a unitary\footnote{As
    $U^{{\rm dis}}_{{\bf k}}$ is a rectangular matrix this is a unitary
    operation in the sense that $(U^{{\rm dis}}_{{\bf
        k}})^{\dagger}U^{{\rm dis}}_{{\bf k}}=1$.}
transformation amongst the Bloch states which fall within the energy window at each kpoint.
 \begin{equation}
| u_{n{\bf k}}^{{\rm opt}}\rangle = \sum_{m\in N^{{\bf k}}_{{\rm win}}}
U^{{\rm dis}}_{mn{\bf k}} | u_{m{\bf k}}\rangle
\end{equation}
where $U^{{\rm dis}}_{{\bf k}}$ is a rectangular $N_{{\rm wann}}\times N^{{\bf
 k}}_{{\rm win}}$
 matrix. The set of $U^{{\rm dis}}_{{\bf k}}$ are obtained by minimising
 the gauge invariant spread, $\Omega_{{\rm I}}$ within the outer energy
 window. The MV procedure can then be used to minimise $\tilde{\Omega}$
 and hence obtain MLWF for this optimal subspace.

It should be noted that the energy bands of this
optimally subspace may not correspond to any of the original energy
bands (due to mixing between states). In order to preserve exactly the
properties of a system in a given energy range (eg around the Fermi
level) we introduce a second  energy window. States lying
within this inner, or ``frozen'' energy window are included unchanged
in the optimal subspace.

\subsection{Citation}
We ask that you acknowledge the use of Wannier90 in any publications
arising from the use of this code through the following reference
\begin{quote}
[ref] A. A. Mostofi, J. R. Yates,
N. Marzari, I. Souza and D. Vanderbilt,  

 http://www.wannier.org/                 
\end{quote}                                                              

It will also be appropriate to cite the original articles:\\\\
{\it Maximally localized generalized Wannier functions for composite energy bands}\\
N. Marzari and D. Vanderbilt, Phys. Rev. B 56, 12847 (1997)\\\\
{\it Maximally localized Wannier functions for entangled energy bands}\\
I. Souza, N. Marzari and D. Vanderbilt, Phys. Rev. B 65, 035109 (2002)


\subsection{Credits}
The present release of Wannier90 was written by Arash Mostofi (Marzari
group @ MIT) and Jonathan Yates (Souza Group @ UCB/LBNL). Wannier90 is
based on routines written in 1996-7 for occupied bands by Nicola Marzari
and David Vanderbilt and for entangled bands by Ivo Souza, Nicola Marzari,
and David Vanderbilt in 2000-1. 

Acknowledgements: ?

Wannier90 \copyright 1997-2006 Jonathan Yates, Arash
Mostofi, Nicola Marzari, Ivo Souza, David Vanderbilt      

\subsection{Licence}
All the material in this distribution is free software; you can
redistribute it and/or 
modify it under the terms of the GNU General Public License
as published by the Free Software Foundation; either version 2
of the License, or (at your option) any later version.

This program is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
GNU General Public License for more details.

You should have received a copy of the GNU General Public License
along with this program; if not, write to the Free Software
Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA  02110-1301, USA.


 
