\section*{About this booklet}
\addcontentsline{toc}{section}{About this booklet}
This solution manual consists of \numexamples{} sections, each containing the solutions, in the form of plots, tabs, and texts, to the corresponding example in the \Wannier{} \version{} tutorial! For each example, only the outline and key questions from the tutorial are reported here.   
 All of the \Wannier{} input files have been provided. From example 5 onwards, input files for the pwscf interface (\url{http://www.
quantum-espresso.org}) to \Wannier{} have also been provided. You will need a recent working version of the \QE{} package (\texttt{v6.2} and above), to run these examples. In particular, you will need \texttt{pw.x} and \texttt{pw2wannier90.x}, as explained in the \Wannier{} \version{} tutorial.
Please visit \url{http://www.quantum-espresso.org} to download the package and follow the instruction on the website for installation. Further details on how to run the calculations for each example may be found in the corresponding section of the \Wannier{} \version{} tutorial.
There are interfaces to a number of other electronic structure codes including: \abinit{} (\url{http://www.
abinit.org}), \fleur{} (\url{http://www.flapw.de}), \OpenMX{} (\url{http://www.openmx-square.org/}), \GPAW{} (\url{https://wiki.fysik.dtu.dk/gpaw/}), \VASP{} (\url{http://www.vasp.at}), and \Wien{} (\url{http://www.wien2k.at}). 

All the tests were performed on an x86\_64 octa-core Intel(R) Xeon(R) CPU E5620 @2.40GHz. Two packages \texttt{intel-suite/2015.3.187} and \texttt{mkl/2015.3.187} were used for the compilation of \Wannier{}. We expect some of the numerical results to depend on the architecture, the compiler distribution, \eg{} \texttt{gcc} \vs{} \texttt{gfortran}, the version of the compiler and the libraries. However, general trends should not be affected by these parameters and, in principle, it should be safe to ignore the differences between different set-ups. If you find that your results are significantly different from the one given in this booklet, please report it to the \Wannier{} developers team by opening an issue on the GitHub repository \url{https://github.com/wannier-developers/wannier90} or by writing an email to the forum at \texttt{wannier@quantum-espresso.org} (we strongly reccommend the first option). Moreover, if you know how to solve the issue and you have a fix for it, you can open a pull-request on the GitHub repo.